### 2022-08-21 chapters/chapter4.tex
update the final version

% Fortunately, a common scenario is goal-oriented RL, where the RL agent is tasked with solving different goals within the same environment~\cite{kaelbling1993learning,schaul2015universal}. Even if each task has a sparse reward, the agent ideally \textit{generalises} across goals, making the learning process easier. For example, in a robotic manipulation task, the goal during a single episode would be to achieve a specific position of a target object.
In the last chapter, we have successfully combined hindsight goal relabelling with a state-of-the-art on-policy RL algorithm and introduced a new goal-oriented RL method -- episodic self-imitation learning (ESIL). Although the experiments show the effectiveness of the proposed method, ESIL still has worse sample efficiency than the vanilla hindsight experience replay (i.e., DDPG+HER). This is because DDPG is an off-policy RL algorithm, and it employs a replay buffer to store transitions during training, which enables the usage of past experiences. In this chapter, we focus on finding the shortcoming of HER and improving it to elevate learning efficiency and performance.

@@

We note that HER, whilst it enables solutions to previously unsolved tasks, can be somewhat inefficient in its use of uniformly sampling transitions during training. In the same way that prioritised experience replay~\cite{schaul2016prioritized} has significantly improved over the standard experience replay in RL, several approaches have improved upon HER by using data-dependent sampling~\cite{fang2019curriculum,zhao2018energy}. HER with energy-based prioritisation (HEBP)~\cite{zhao2018energy} assumes semantic knowledge about the goal-space and uses the energy of the target objects to sample trajectories with high energies, and then samples transitions uniformly. Curriculum-guided HER (CHER)~\cite{fang2019curriculum} samples trajectories uniformly, and then samples transitions based on a mixture of proximity to the desired goal and the diversity of the samples; CHER adapts the weighting of these factors over time. In this chapter, we introduce diversity-based trajectory and goal selection with HER (DTGSH; See Figure~\ref{fig:illustration_c4}), which samples trajectories based on the diversity of the goals achieved within the trajectory, and then samples transitions based on the diversity of the set of samples. In the experiments, DTGSH is evaluated on five challenging robot manipulation tasks. From extensive experiments, our proposed method converges faster and reaches higher rewards than prior work, without requiring domain knowledge~\cite{zhao2018energy} or tuning a curriculum~\cite{fang2019curriculum}, and is based on a single concept---determinantal point processes (DPPs)~\cite{kulesza2012determinantal}.
\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{figures/chapter4/illustration_DTGSH_latest.pdf}
    \caption{Overview of DTGSH. Every time a new episode is completed, its diversity is calculated, and it is stored in the episodic replay buffer. During training, $m$ episodes are sampled according to their diversity-based priority, and then $k$ diverse, hindsight-relabelled transitions are sampled using a $k$-DPP~\cite{kulesza2011k}.}
    \label{fig:illustration_c4}
\end{figure}

@@

@@

In our approach, DPPs are used to model the diversity of achieved goal states $g^{ac}_{t}$ in an episode, or subsets thereof. For a single trajectory $\Tau$ of length $T$, we divide it into several partial trajectories $\tau_{j}$ of length $b$, with achieved goal states $\{g^{ac}_{t}\}_{t=n:n+b-1}$. That is, with a sliding window of $b = 2$, a trajectory $\Tau$ can be divided into $N_p$ partial trajectories:
\begin{equation}
    \Tau_{i} = \{\{\underbrace{g^{ac}_{1}, g^{ac}_{2}}_{\tau_{1}}\}, \{\underbrace{g^{ac}_{2}, g^{ac}_{3}}_{\tau_{2}}\}, \{\underbrace{g^{ac}_{3}, g^{ac}_{4}}_{\tau_{3}}\}, \cdots, \{\underbrace{g^{ac}_{T-1}, g^{ac}_{T}}_{\tau_{N_{p}}}\}\}.
\end{equation}
The diversity $d_{\tau_{j}}$ of each partial trajectory $\tau_{j}$ can be computed as:
\begin{equation}
    d_{\tau_{j}} = \text{det}(\mathbf{L}_{\tau_{j}}), 
\label{eq:partial_div}
\end{equation}
where $\mathbf{L}_{\tau_{j}}$ is the kernel matrix of partial trajectory $\tau_{j}$:
\begin{equation}
    L_{\tau_{j}} = \mathbf{M}^{T}\mathbf{M},
\end{equation}
and $\mathbf{M}=[\hat{g}^{ac}_{n}, \hat{g}^{ac}_{n+1}, \cdots, \hat{g}^{ac}_{n+b-1}]$, where each $\hat{g}^{ac}$ is the $\ell_2$-normalised version of the achieved goal $g^{ac}$~\cite{kulesza2011k}. Finally, the diversity $d_\Tau$ of trajectory $\Tau$ is the sum of the diversity of its $N_p$ constituent partial trajectories:
\begin{equation}
    d_\Tau = \sum_{j=1}^{N_{p}} d_{\tau_{j}}.
\label{eq:total_div}
\end{equation}

@@

\subsection{Diversity-based Goal Selection}
In prior work~\cite{andrychowicz2017hindsight,zhao2018energy}, after selecting the trajectories from the replay buffer, one transition from each selected trajectory is sampled uniformly to construct a minibatch for training. However, the modified goals $g^{\prime}$ in the minibatch might be similar, resulting in redundant information. In order to form a minibatch with diverse goals for more efficient learning, we use $k$-DPPs~\cite{kulesza2011k} for sampling goals. Compared to the standard DPP, a $k$-DPP is a conditional DPP where the subset $Y$ has a fixed size $k$, with the probability distribution function:
\begin{equation}
    \mathcal{P}_{L}^{k}(\mathbf{Y}=Y) = \frac{\text{det}(\mathbf{L}_{Y})}{\sum_{|Y^{\prime}|=k} \text{det}(\mathbf{L}_{Y^{\prime}})}.
\end{equation}
$k$-DPPs are more appropriate for goal selection with a minibatch of fixed size $k$. Given $m > k$ trajectories sampled from the replay buffer, we first uniformly sample a transition from each of the $m$ trajectories. Finally, a $k$-DPP is used to sample a diverse set of transitions based on the relabelled goals $g'$ (which, in this context, we denote as ``candidate goals''). Figure~\ref{fig:goal_sampling} gives an example of uniform vs. $k$-DPP sampling, demonstrating the increased coverage of the latter. Figure~\ref{fig:kde} provides corresponding estimated density plots; note that the density of the $k$-DPP samples is actually more uniform over the \textit{support} of the candidate goal distribution.
\begin{figure}[h]
    \begin{subfigure}{\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/chapter4/goals_distribution_2d.pdf}
        \caption{Plot of candidate goals and selected goals. $k$-DPP sampling is more likely to sample points from the full span of the goal space.}
        \label{fig:goal_sampling}
    \end{subfigure}
    \begin{subfigure}{\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/chapter4/kde.pdf}
        \caption{Kernel density estimation of the distributions of goals. $k$-DPP leads to a more uniform selection of goals over the support of the goal space.}
        \label{fig:kde}
    \end{subfigure}
    \caption{Visualisation of $k=32$ goals selected from $m=100$ candidate goals of the Push task using either uniform sampling or $k$-DPP sampling, respectively. The candidate goals are distributed over a 2D ($xy$) space. Note that  $k$-DPP sampling (right hand plots) results in a broader span of selected goals in $xy$ space compared to uniform sampling (left hand plots).}
\end{figure}

@@

\begin{figure}
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/push_resize.png}
    \caption{Push}
    \label{subfig:env_push}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/pick_resize.png}
    \caption{Pick$\&$Place}
    \label{subfig:env_pick}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/handegg_resize.png}
    \caption{EggFull}
    \label{subfig:env_handegg}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/handblock_resize.png}
    \caption{BlockRotate}
    \label{subfig:env_handblock}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/handpen_resize.png}
    \caption{PenRotate}
    \label{subfig:env_handpen}
  \end{subfigure}\hfill
  \caption{Robot manipulation environments. (a-b) use the Fetch robot, and (c-e) use the Shadow Dexterous Hand.} 
  \label{fig:env_c4}
\end{figure}

@@

\begin{figure}
\centering
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/FetchPush-v1.pdf}
    \caption{Push}
    \label{subfig:baseline_push}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/FetchPickAndPlace-v1.pdf}
    \caption{Pick\&Place}
    \label{subfig:baseline_pick}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateEggFull-v0.pdf}
    \caption{EggFull}
    \label{subfig:baseline_handegg}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateBlockRotateXYZ-v0.pdf}
    \caption{BlockRotate}
    \label{subfig:baseline_handblock}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulatePenRotate-v0.pdf}
    \caption{PenRotate}
    \label{subfig:baseline_handpen}
  \end{subfigure}\hfill
%   \begin{subfigure}[t]{0.49\textwidth}
%     \includegraphics[width=\textwidth]{figures/chapter4/blank.png}
%   \end{subfigure}\hfill
  \caption{Success rate of DTGSH and baseline approaches. \td{DDPG+DTGSH outpeforms other baselines in all tasks, and without using HER, DDGP cannot even complete the simple Reach task.}} 
  \label{fig:baseline_compare_c4}
\end{figure}

Figure~\ref{fig:baseline_compare_c4} and Table~\ref{tab:benchmark} show the performance of DDPG+DTGSH and baseline approaches on all five tasks. In the Fetch tasks, DDPG+DTGSH and DDPG+HEBP both learn significantly faster than the other methods, while in the Shadow Dexterous Hand tasks DDPG+DTGSH learns the fastest and achieves higher success rates than all other methods. In particular, DDPG cannot solve any tasks without using HER, and CHER performs worse in the Fetch tasks. We believe the results highlight the importance of sampling both diverse trajectories \textit{and} goals, as in our proposed method, DTGSH.
\begin{table}[h]
    \centering
    \resizebox{\textwidth}{!}{
    \begin{tabular}{c|c@{\hskip 0.1in}|c@{\hskip 0.1in}|c|c|c}
    \toprule
         & Push & Pick\&Place & EggFull & BlockRotate & PenRotate \\
        \midrule
        DDPG & 0.09$\pm$0.01 & 0.04$\pm$0.00 & 0.00$\pm$0.00 & 0.01$\pm$0.00 & 0.00$\pm$0.00 \\
        DDPG+HER & \textbf{1.00$\pm$0.00} & 0.89$\pm$0.03 & 0.11$\pm$0.01 & 0.55$\pm$0.04 & 0.15$\pm$0.02 \\
        DDPG+HEBP & \textbf{1.00$\pm$0.00} & 0.91$\pm$0.03 & 0.14$\pm$0.02 & 0.59$\pm$0.02 & 0.20$\pm$0.03\\
        DDPG+CHER & \textbf{1.00$\pm$0.00} & 0.91$\pm$0.04 & 0.15$\pm$0.01 & 0.54$\pm$0.04 & 0.17$\pm$0.03 \\
        \midrule
        DDPG+DTGSH & \textbf{1.00$\pm$0.00} & \textbf{0.94$\pm$0.01} & \textbf{0.17$\pm$0.03} & \textbf{0.62$\pm$0.02} & \textbf{0.21$\pm$0.02}\\
        \bottomrule
    \end{tabular}
    }
    \vspace{0.2em}
    \caption{Final mean success rate $\pm$ standard deviation, with best results in \textbf{bold}.}
    %\vspace{-14mm}
    \label{tab:benchmark}
\end{table}
%\vspace{-15mm}
\subsection{Results on Ablation Studies}
In this section, we perform the following experiments to investigate the effectiveness of each component in DTGSH: 1) diversity-based trajectory selection with HER (DTSH) and diversity-based goal selection with HER (DGSH) are evaluated independently to assess the contribution of each stage, 2) the performance using different partial trajectory lengths $b$, and 3) the performance of using different candidate goal set sizes $m$.
\begin{figure}[h!]
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/FetchPush-v1_ab1.pdf}
    \caption{Push}
    \label{subfig:baseline_push_ab1}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/FetchPickAndPlace-v1_ab1.pdf}
    \caption{Pick\&Place}
    \label{subfig:baseline_pick_ab1}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateEggFull-v0_ab1.pdf}
    \caption{EggFull}
    \label{subfig:baseline_handegg_ab1}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateBlockRotateXYZ-v0_ab1.pdf}
    \caption{BlockRotate}
    \label{subfig:baseline_handblock_ab1}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulatePenRotate-v0_ab1.pdf}
    \caption{PenRotate}
    \label{subfig:baseline_handpen_ab1}
  \end{subfigure}\hfill
%   \begin{subfigure}[t]{0.33\textwidth}
%     \includegraphics[width=\textwidth]{figures/chapter4/blank.png}
%   \end{subfigure}\hfill
  \caption{Success rate of HER, DTGSH, and ablations DTSH and DGSH. \td{From the results, DTSH outperforms DDPG+HER in all tasks, which verify the efficiency that selecting trajectories with diverse achieved goals for training. DGSH also has better performance than DDPG+HER in three tasks. By combining these two modules together, DTGSH achieves the best performance.} } 
  \label{fig:ablation1}
\end{figure}

@@

\subsection{Time Complexity}
Table~\ref{tab:time} gives example training times of all of the HER-based algorithms. DTGSH requires an additional calculation of the diversity score of $\mathcal{O}(N_{p}b^3)$ at the end of every training episode, and sampling of $\mathcal{O}(mk^2)$ for each minibatch.
\begin{table}[h!]
    \centering
    \resizebox{0.85\textwidth}{!}{
    \begin{tabular}{c|c|c|c|c}
    \toprule
           & DDPG+HER & DDPG+HEBP & DDPG+CHER & DDPG+DTGSH\\
        \midrule
        Time  & 00:55:08 & 00:56:32 & 03:02:18 & 01:52:30 \\
        \bottomrule
    \end{tabular}
    }
    \vspace{0.2em}
    \caption{Training time (hours:minutes:seconds) of DTGSH and baseline approaches on the Push task for 50 epochs.}
    \label{tab:time}
\end{table}

>>>>

% Fortunately, a common scenario is goal-oriented RL, where the RL agent is tasked with solving different goals within the same environment~\cite{kaelbling1993learning,schaul2015universal}. Even if each task has a sparse reward, the agent ideally \textit{generalises} across goals, making the learning process easier. For example, in a robotic manipulation task, the goal during a single episode would be to achieve a specific position of a target object.
In the last chapter, we successfully combined hindsight goal relabelling with a state-of-the-art on-policy RL algorithm and introduced a new goal-oriented RL method -- episodic self-imitation learning (ESIL). In ESIL, a novel self-imitation learning loss is used to accelerate the training by using the hindsight experience. Although the experiments show the effectiveness of the proposed method, ESIL still has worse sample efficiency than the vanilla hindsight experience replay (i.e., DDPG+HER). This is because DDPG is an off-policy RL algorithm, and it employs a replay buffer to store transitions during training, which enables the usage of past experiences. In this chapter, we focus on finding the shortcoming of HER and improving it to elevate learning efficiency and performance by proposing a new sampling strategy.

@@

We note that HER, whilst it enables solutions to previously unsolved tasks, can be somewhat inefficient in its use of uniformly sampling transitions during training. In the same way that prioritised experience replay~\cite{schaul2016prioritized} has significantly improved over the standard experience replay in RL, several approaches have improved upon HER by using data-dependent sampling~\cite{fang2019curriculum,zhao2018energy}. HER with energy-based prioritisation (HEBP)~\cite{zhao2018energy} assumes semantic knowledge about the goal-space and uses the energy of the target objects to sample trajectories with high energies, and then samples transitions uniformly. Curriculum-guided HER (CHER)~\cite{fang2019curriculum} samples trajectories uniformly, and then samples transitions based on a mixture of proximity to the desired goal and the diversity of the samples; CHER adapts the weighting of these factors over time. In this chapter, we introduce diversity-based trajectory and goal selection with HER (DTGSH; See Figure~\ref{fig:illustration_c4}), which samples trajectories based on the diversity of the goals achieved within the trajectory, and then samples transitions based on the diversity of the set of samples. In the experiments, DTGSH is evaluated on five challenging robot manipulation tasks. From extensive experiments, our proposed method converges faster and reaches higher rewards than prior work, without requiring domain knowledge~\cite{zhao2018energy} or tuning a curriculum~\cite{fang2019curriculum}, and is based on a single concept---determinantal point processes (DPPs)~\cite{kulesza2012determinantal}.
\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{figures/chapter4/illustration_DTGSH_latest.pdf}
    \caption[Overview of DTGSH.]{Overview of DTGSH. Every time a new episode is completed, its diversity is calculated, and it is stored in the episodic replay buffer. During training, $m$ episodes are sampled according to their diversity-based priority, and then $k$ diverse, hindsight-relabelled transitions are sampled using a $k$-DPP~\cite{kulesza2011k}.}
    \label{fig:illustration_c4}
\end{figure}

@@

% \begin{table}[h]
%     \centering
%     \resizebox{\textwidth}{!}{
%     \begin{tabular}{c|c@{\hskip 0.1in}|c@{\hskip 0.1in}|c|c}
%     \toprule
%          & DDPG+HER~\cite{andrychowicz2017hindsight} & DDPG+HEBP~\cite{zhao2018energy} & DDPG+CHER~\cite{fang2019curriculum} & DDPG+DTGSH (Ours) \\
%         \midrule
%         Sampling Strategy (Trajectory) & & & & \\
%         Sampling Strategy (Transition) & & & & \\
%         \bottomrule
%     \end{tabular}
%     }
%     \vspace{0.2em}
%     \caption[Average success rate of DTGSH and other baselines in the final epoch during training.]{Final mean success rate $\pm$ standard deviation, with best results in \textbf{bold}.}
%     %\vspace{-14mm}
%     \label{tab:benchmark}
% \end{table}

\begin{table}[h]
\resizebox{\textwidth}{!}{
\begin{tabular}{|c|c|c|c|}
\hline
& DDPG+HER~\cite{andrychowicz2017hindsight} & DDPG+HEBP~\cite{zhao2018energy} & DDPG+CHER~\cite{fang2019curriculum} \\ \hline
\begin{tabular}[c]{@{}c@{}}Sampling Strategy\\       (Trajectory)\end{tabular} & uniform  & \begin{tabular}[c]{@{}c@{}}according to the energy\\ of each trajectory\end{tabular} & uniform                                                                                                \\ \hline
\begin{tabular}[c]{@{}c@{}}Sampling Strategy\\ (Transition)\end{tabular}       & uniform  & uniform                                                                              & \begin{tabular}[c]{@{}c@{}}according to the proximity and \\ diversity of candidate goals\end{tabular} \\ \hline
\end{tabular}
}
\caption[Comparison of sampling strategies between different HER related methods.]{The comparison of sampling strategies between HER related methods.}
\label{tab:sample_strategy}
\end{table}

We further use Table~\ref{tab:sample_strategy} to summarize the sampling strategies of baseline HER related algorithms in this chapter. Although HEBP~\cite{zhao2018energy} and CHER~\cite{fang2019curriculum} improve the sample efficiency of HER by proposing new sampling strategies, these methods still use uniform sampling strategy for either trajectory or transition selection, which may sample redundant trajectories and transitions, and thus limit the performance. In contrast to these methods, in DTGSH, the sampling strategies for both trajectory and transition are improved, which increase the chance to sample trajectories and transitions that contribute more to the training.

@@

In our approach, DPPs are used to model the diversity of achieved goal states $g^{ac}_{t}$ in an episode, or subsets thereof. For a single trajectory $\Tau$ of length $T$, we divide it into several partial trajectories $\tau_{j}$ of length $b$, with achieved goal states $\{g^{ac}_{t}\}_{t=n:n+b-1}$. That is, with a sliding window of $b = 2$, a trajectory $\Tau$ can be divided into $N_p$ partial trajectories:
\begin{equation}
    \Tau_{i} = \{\{\underbrace{g^{ac}_{1}, g^{ac}_{2}}_{\tau_{1}}\}, \{\underbrace{g^{ac}_{2}, g^{ac}_{3}}_{\tau_{2}}\}, \{\underbrace{g^{ac}_{3}, g^{ac}_{4}}_{\tau_{3}}\}, \cdots, \{\underbrace{g^{ac}_{T-1}, g^{ac}_{T}}_{\tau_{N_{p}}}\}\}.
\end{equation}
The diversity $d_{\tau_{j}}$ of each partial trajectory $\tau_{j}$ can be computed as:
\begin{equation}
    d_{\tau_{j}} = \text{det}(\mathbf{L}_{\tau_{j}}), 
\label{eq:partial_div}
\end{equation}
where $\mathbf{L}_{\tau_{j}}$ is the kernel matrix of partial trajectory $\tau_{j}$:
\begin{equation}
    L_{\tau_{j}} = \mathbf{M}^{T}\mathbf{M},
\end{equation}
and $\mathbf{M}=[\hat{g}^{ac}_{n}, \hat{g}^{ac}_{n+1}, \cdots, \hat{g}^{ac}_{n+b-1}]$, where each $\hat{g}^{ac}$ is the $\ell_2$-normalised vector of the achieved goal $g^{ac}$~\cite{kulesza2011k}. Finally, the diversity $d_\Tau$ of trajectory $\Tau$ is the sum of the diversity of its $N_p$ constituent partial trajectories:
\begin{equation}
    d_\Tau = \sum_{j=1}^{N_{p}} d_{\tau_{j}}.
\label{eq:total_div}
\end{equation}

@@

\subsection{Diversity-based Goal Selection}
In prior work~\cite{andrychowicz2017hindsight,zhao2018energy}, after selecting the trajectories from the replay buffer, one transition from each selected trajectory is sampled uniformly to construct a minibatch for training. However, the modified goals $g^{\prime}$ in the minibatch might be similar, resulting in redundant information. In order to form a minibatch with diverse goals for more efficient learning, we use $k$-DPPs~\cite{kulesza2011k} for sampling goals. Compared to the standard DPP, a $k$-DPP is a conditional DPP where the subset $Y$ has a fixed size $k$, with the probability distribution function:
\begin{equation}
    \mathcal{P}_{L}^{k}(\mathbf{Y}=Y) = \frac{\text{det}(\mathbf{L}_{Y})}{\sum_{|Y^{\prime}|=k} \text{det}(\mathbf{L}_{Y^{\prime}})}.
\end{equation}
$k$-DPPs are more appropriate for goal selection with a minibatch of fixed size $k$. Given $m > k$ trajectories sampled from the replay buffer, we first uniformly sample a transition from each of the $m$ trajectories. Finally, a $k$-DPP is used to sample a diverse set of transitions based on the relabelled goals $g'$ (which, in this context, we denote as ``candidate goals''). Figure~\ref{fig:goal_sampling} gives an example of uniform vs. $k$-DPP sampling, demonstrating the increased coverage of the latter. Figure~\ref{fig:kde} provides corresponding estimated density plots; note that the density of the $k$-DPP samples is actually more uniform over the \textit{support} of the candidate goal distribution.
\begin{figure}[h]
    \begin{subfigure}{\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/chapter4/goals_distribution_2d.pdf}
        \caption{Plot of candidate goals and selected goals. $k$-DPP sampling is more likely to sample points from the full span of the goal space.}
        \label{fig:goal_sampling}
    \end{subfigure}
    \begin{subfigure}{\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/chapter4/kde.pdf}
        \caption{Kernel density estimation of the distributions of goals. $k$-DPP leads to a more uniform selection of goals over the support of the goal space.}
        \label{fig:kde}
    \end{subfigure}
    \caption[Visualisation of goal selection between uniform sampling and $k$-DPP sampling.]{Visualisation of $k=32$ goals selected from $m=100$ candidate goals of the Push task using either uniform sampling or $k$-DPP sampling, respectively. The candidate goals are distributed over a 2D ($xy$) space. Note that  $k$-DPP sampling (right hand plots) results in a broader span of selected goals in $xy$ space compared to uniform sampling (left hand plots).}
\end{figure}

@@

\begin{figure}
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/push_resize.png}
    \caption{Push}
    \label{subfig:env_push}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/pick_resize.png}
    \caption{Pick$\&$Place}
    \label{subfig:env_pick}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/handegg_resize.png}
    \caption{EggFull}
    \label{subfig:env_handegg}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/handblock_resize.png}
    \caption{BlockRotate}
    \label{subfig:env_handblock}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/handpen_resize.png}
    \caption{PenRotate}
    \label{subfig:env_handpen}
  \end{subfigure}\hfill
  \caption[Evaluation environments of DTGSH.]{Robot manipulation environments. (a-b) use the Fetch robot, and (c-e) use the Shadow Dexterous Hand.} 
  \label{fig:env_c4}
\end{figure}

@@

\begin{figure}[H]
\centering
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/FetchPush-v1.pdf}
    \caption{Push}
    \label{subfig:baseline_push}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/FetchPickAndPlace-v1.pdf}
    \caption{Pick\&Place}
    \label{subfig:baseline_pick}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateEggFull-v0.pdf}
    \caption{EggFull}
    \label{subfig:baseline_handegg}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateBlockRotateXYZ-v0.pdf}
    \caption{BlockRotate}
    \label{subfig:baseline_handblock}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulatePenRotate-v0.pdf}
    \caption{PenRotate}
    \label{subfig:baseline_handpen}
  \end{subfigure}\hfill
%   \begin{subfigure}[t]{0.49\textwidth}
%     \includegraphics[width=\textwidth]{figures/chapter4/blank.png}
%   \end{subfigure}\hfill
  \caption[Results of DTGSH in the robotic manipulation environment.]{Success rate of DTGSH and baseline approaches. DDPG+DTGSH outperforms other baselines in all tasks, and without using HER, DDGP cannot even complete the simple Reach task.}
  \label{fig:baseline_compare_c4}
\end{figure}

Figure~\ref{fig:baseline_compare_c4} and Table~\ref{tab:benchmark} show the performance of DDPG+DTGSH and baseline approaches on all five tasks. In the Fetch tasks, DDPG+DTGSH and DDPG+HEBP both learn significantly faster than the other methods, while in the Shadow Dexterous Hand tasks DDPG+DTGSH learns the fastest and achieves higher success rates than all other methods. In particular, DDPG cannot solve any tasks without using HER, and CHER performs worse in the Fetch tasks. We believe the results highlight the importance of sampling both diverse trajectories \textit{and} goals, as in our proposed method, DTGSH.
\begin{table}[h]
    \centering
    \resizebox{\textwidth}{!}{
    \begin{tabular}{c|c@{\hskip 0.1in}|c@{\hskip 0.1in}|c|c|c}
    \toprule
         & Push & Pick\&Place & EggFull & BlockRotate & PenRotate \\
        \midrule
        DDPG & 0.09$\pm$0.01 & 0.04$\pm$0.00 & 0.00$\pm$0.00 & 0.01$\pm$0.00 & 0.00$\pm$0.00 \\
        DDPG+HER & \textbf{1.00$\pm$0.00} & 0.89$\pm$0.03 & 0.11$\pm$0.01 & 0.55$\pm$0.04 & 0.15$\pm$0.02 \\
        DDPG+HEBP & \textbf{1.00$\pm$0.00} & 0.91$\pm$0.03 & 0.14$\pm$0.02 & 0.59$\pm$0.02 & 0.20$\pm$0.03\\
        DDPG+CHER & \textbf{1.00$\pm$0.00} & 0.91$\pm$0.04 & 0.15$\pm$0.01 & 0.54$\pm$0.04 & 0.17$\pm$0.03 \\
        \midrule
        DDPG+DTGSH & \textbf{1.00$\pm$0.00} & \textbf{0.94$\pm$0.01} & \textbf{0.17$\pm$0.03} & \textbf{0.62$\pm$0.02} & \textbf{0.21$\pm$0.02}\\
        \bottomrule
    \end{tabular}
    }
    \vspace{0.2em}
    \caption[Average success rate of DTGSH and other baselines in the final epoch during training.]{Final mean success rate $\pm$ standard deviation, with best results in \textbf{bold}.}
    %\vspace{-14mm}
    \label{tab:benchmark}
\end{table}
%\vspace{-15mm}
\subsection{Results on Ablation Studies}
In this section, we perform the following experiments to investigate the effectiveness of each component in DTGSH: 1) diversity-based trajectory selection with HER (DTSH) and diversity-based goal selection with HER (DGSH) are evaluated independently to assess the contribution of each stage, 2) the performance using different partial trajectory lengths $b$, and 3) the performance of using different candidate goal set sizes $m$.
\begin{figure}[h!]
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/FetchPush-v1_ab1.pdf}
    \caption{Push}
    \label{subfig:baseline_push_ab1}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/FetchPickAndPlace-v1_ab1.pdf}
    \caption{Pick\&Place}
    \label{subfig:baseline_pick_ab1}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateEggFull-v0_ab1.pdf}
    \caption{EggFull}
    \label{subfig:baseline_handegg_ab1}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulateBlockRotateXYZ-v0_ab1.pdf}
    \caption{BlockRotate}
    \label{subfig:baseline_handblock_ab1}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/HandManipulatePenRotate-v0_ab1.pdf}
    \caption{PenRotate}
    \label{subfig:baseline_handpen_ab1}
  \end{subfigure}\hfill
%   \begin{subfigure}[t]{0.33\textwidth}
%     \includegraphics[width=\textwidth]{figures/chapter4/blank.png}
%   \end{subfigure}\hfill
  \caption[Results of ablation studies of different modules.]{Success rate of HER, DTGSH, and ablations DTSH and DGSH. From the results, DTSH outperforms DDPG+HER in all tasks, which verify the efficiency that selecting trajectories with diverse achieved goals for training. DGSH also has better performance than DDPG+HER in three tasks. By combining these two modules together, DTGSH achieves the best performance.}
  \label{fig:ablation1}
\end{figure}

@@

%\subsection{Time Complexity}
% Table~\ref{tab:time} gives example training times of all of the HER-based algorithms. DTGSH requires an additional calculation of the diversity score of $\mathcal{O}(N_{p}b^3)$ at the end of every training episode, and sampling of $\mathcal{O}(mk^2)$ for each minibatch.

\section{Discussion}
In this chapter, we introduced diversity-based trajectory and goal selection with hindsight experience replay (DTGSH) to improve the learning efficiency of goal-orientated RL agents in sparse reward settings. Our method can be divided into two stages: 1) valuable trajectories are selected according to diversity-based priority, as modelled by determinantal point processes (DPPs)~\cite{kulesza2012determinantal}, and 2) $k$-DPPs~\cite{kulesza2011k} are leveraged to sample transitions with diverse goal states from previously selected trajectories for training. Our experiments empirically show that DTGSH achieves faster learning and higher final performance in five challenging robot manipulation tasks, compared to previous state-of-the-art approaches~\cite{andrychowicz2017hindsight,fang2019curriculum,zhao2018energy}. In addition, unlike prior extensions of hindsight experience replay, DTGSH does not require semantic knowledge of the goal space~\cite{zhao2018energy}, and does not require tuning a curriculum~\cite{fang2019curriculum}.

\begin{table}[h!]
    \centering
    \resizebox{0.85\textwidth}{!}{
    \begin{tabular}{c|c|c|c|c}
    \toprule
           & DDPG+HER & DDPG+HEBP & DDPG+CHER & DDPG+DTGSH\\
        \midrule
        Time  & 00:55:08 & 00:56:32 & 03:02:18 & 01:52:30 \\
        \bottomrule
    \end{tabular}
    }
    \vspace{0.2em}
    \caption[Training time comparison between DTGSH and other baselines.]{Training time (hours:minutes:seconds) of DTGSH and baseline approaches on the Push task for 50 epochs.}
    \label{tab:time}
\end{table}
